A sheaf--function dictionary for algebraic tori over non-archimedean local fields

### 2013-08-16 shanghai.tex
Working on style in sections 1 and 2

As mentioned after \cite{deligne:80a}*{Def.~1.1.10}, a Weil sheaf $(\gqcs{L},\phi)$ on $G$ may be interpreted as a constructible $\ell$-adic sheaf $\gqcs{L}$ on $\bG$ together with an action of the Weil group $\Weil{\Fq}$ on $\gqcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ on $\bG$. We now review this point of view for later use in Section \ref{sec:basechange}.  If $w = \Frob{\Fq}^n$ we write $w_G$ for $\Frob{G}^n$.  For each such $w$ we define an isomorphism $\varphi(w) : w_G^* \gqcs{L}\to \gqcs{L}$ by
\[
\varphi(w) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi);
\]
these satisfy $\varphi(uv) = \varphi(v) \circ v_G^* \varphi(u)$ for $u,v\in \Weil{\Fq}$, and
we have $\phi = \varphi(\Frob{\Fq})$. 
If we further define $\varphi_1(w) \ceq (w_G)_*(\varphi(w)^{-1})$ then
$\varphi_1(w) : (w_G)_* \gqcs{L}\to \gqcs{L}$ is an isomorphism and
$\varphi_1(uv) = \varphi_1(u) \circ (u_G)_* \varphi_1(v)$ for $u,v\in \Weil{\Fq}$.
Thus the pair $(\gqcs{L},\varphi_1)$ almost satisfies the criteria for an action of $\Weil{\Fq}$ on $\gqcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of \cite{deligne:SGA4.5}*{Expos\'e XIII,~1.1}, failing only because $\Weil{\Fq}$ is profinite \David{Do you mean to say that $\Weil{\Fq}$ is not profinite?}. This fact \David{Which fact?  That we can think about quasicharacter sheaves as an action on $\gqcs{L}$?  Or that they fail to fulfill the requirements?} goes to the heart of the notion of a Weil sheaf and is one of the essential aspects of our definition of $\QC(G)$. 
%\end{remark}

@@

%\begin{proposition}\label{tensor}
The category $\QC(G)$ of quasicharacter sheaves on $G$ is rigid symmetric monoidal category \cite{etingof:09a} under the tensor product
$\qcs{L} \otimes_{\EE}\qcs{L'} \ceq (\gqcs{L}\otimes\gqcs{L'}, \mu\otimes\mu', \phi\otimes_{\EE}\phi')$
\David{Why only some $\otimes_{\EE}$?}
with duals given by applying the functor 
$\sheafHom(\ - \ ,\EE)$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\QC(G)$
is not a tensor category in the sense of \cite{deligne:02a}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\QC(G)$ puts a group
structure on the set $\QCiso{G}$ of isomorphism classes in
$\QC(G)$. Our main result regarding quasicharacter sheaves on $G$ (Theorem~\ref{thm:snake})
determines $\QCiso{G}$ under the hypothesis that the component group scheme for $G$ is geometrically finitely generated; see Theorem~\ref{thm:snake}. To do this, we use the trace of Frobenius.

>>>>

As mentioned after \cite{deligne:80a}*{Def.~1.1.10}, a Weil sheaf $(\gqcs{L},\phi)$ on $G$ may be interpreted as a constructible $\ell$-adic sheaf $\gqcs{L}$ on $\bG$ together with an action of the Weil group $\Weil{\Fq}$ on $\gqcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ on $\bG$. We now review this point of view for later use in Section \ref{sec:basechange}.  If $w = \Frob{\Fq}^n$ we write $w_G$ for $\Frob{G}^n$.  For each such $w$ we define an isomorphism $\varphi(w) : w_G^* \gqcs{L}\to \gqcs{L}$ by
\[
\varphi(w) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi);
\]
these satisfy $\varphi(uv) = \varphi(v) \circ v_G^* \varphi(u)$ for $u,v\in \Weil{\Fq}$, and
we have $\phi = \varphi(\Frob{\Fq})$. 
If we further define $\varphi_1(w) \ceq (w_G)_*(\varphi(w)^{-1})$ then
$\varphi_1(w) : (w_G)_* \gqcs{L}\to \gqcs{L}$ is an isomorphism and
$\varphi_1(uv) = \varphi_1(u) \circ (u_G)_* \varphi_1(v)$ for $u,v\in \Weil{\Fq}$.
Thus the pair $(\gqcs{L},\varphi_1)$ almost satisfies the criteria for an action of $\Weil{\Fq}$ on $\gqcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of \cite{deligne:SGA4.5}*{Expos\'e XIII,~1.1}, failing only because $\Weil{\Fq}$ is not profinite. Our use of $\Weil{\Fq}$ rather than $\Gal(\bFq/\Fq)$ allows quasicharacter sheaves to correspond to unbounded characters.
%\end{remark}

@@

The category $\QC(G)$ of quasicharacter sheaves on $G$ is a rigid symmetric monoidal category
\cite{etingof:09a} under the tensor product
$\qcs{L} \otimes \qcs{L'} \ceq (\gqcs{L}\otimes\gqcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$,
with duals given by applying the sheaf-hom functor 
$\sheafHom(\ - \ ,\EE)$.
The category of characters sheaves on $G$ is not abelian, so $\QC(G)$
is not a tensor category in the sense of \cite{deligne:02a}.
This rigid monoidal category structure for $\QC(G)$ puts a group
structure on the set $\QCiso{G}$ of isomorphism classes in $\QC(G)$.
We now begin developing the tools to relate $\QCiso{G}$ to $\Hom(G(\Fq), \EEx)$.
